Applying WaveNet on frequency data

### 2017-01-20 paper/paper.tex
Added section "results"

\begin{document}
\title{Applying WaveNet on frequency data}

@@

\section{Results}
The generated audio files start with some fastly changing frequencies and then after 30-40 frames (1-2 seconds) it stabilizes on one distribution of frequencies. From then on the values of the frequencies do not change any more. Figures \ref{fig:original_frequencies} and \ref{fig:generated_frequencies} show that the neural network is not able to produce a change of frequencies while a real piece of music observes a change every 20-200 frames.\\
The consequence of these frequencies not changing is, that the audio file consists after a noise at the beginning only of a steady pitch. And it doesn't resemble a pitch either.\\

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\columnwidth]{figures/original_frequencies.png}
  \caption{Frequencies of a piano piece used for training}
  \label{fig:original_frequencies}
\end{figure}
\begin{figure}[tbp]
  \centering
  \includegraphics[width=\columnwidth]{figures/generated_frequencies_20s.png}
  \caption{Frequencies of the generated audio piece}
  \label{fig:generated_frequencies}
\end{figure}

>>>>

\pagestyle{plain}
\begin{document}
\title{Applying WaveNet on frequency data}

@@

\section{Results}
The results of all experiments are summarized in a python notebook of the GitHub repository.\footnote{"evaluate\_results.ipynb", available in the subfolder "notebooks"} Unfortunately, the adopted network could not perform as well as the original WaveNet architecture. All generated audio files stabilized on one distribution of frequencies after a few seconds. From then on the values of the frequencies did not change any more. Instead of learning the complete range of frequencies, which define the features of the audio input, the network could only learn some low basis frequencies being constantly present all the time.\\
Figures \ref{fig:original_frequencies} and \ref{fig:generated_frequencies} compare the frequencies of a recorded audio piece from the input data and the frequencies of a generated sample. In the first 200 time frames, the frequency distributions look quite similar. However, afterwards the neural network is not able to produce frequency changes continuously, while a real piece of music observes a change every 20-60 frames.\\
Furthermore the original and generated audios differ in their frequency range. Whereas the original music piece only consists of frequencies up to 32, the generated audio contains frequencies up to 105. This shows that the network has difficulties to decide which frequencies are important and cannot produce as compact and compressed audio output as it is achieved by using FFT and PCA on recorded audio data.\\
Nevertheless, the experiments of this paper can be analysed more detailled. Varying the stride and including normalization yielded conceptually different results.\\
Overlapping chunks of input data turned out to be essential for learning audio features properly. With a stride of 512 (resulting in an overlap of 75\% of neighboring data chunks) results with dynamic volume changes and frequencies within the range $[0, 105]$ within the first seconds of the generated audio could be achieved. In contrast to that, with a stride of 2048 (meaning no overlap) the network only grasped really frequencies in the range $[0,1.6]$ producing no volume or frequency dynamic in the generated audio at all.\textcolor{red}{Missing for discussion: on the one hand overlap necessary, on the other hand emphasises this to learn constant tones.}\\
A second conceptual attempt during the experiments was to normalize the PCA results before the training. The aim of this additional preprocessing task was to balance the contribution of all 100 PCA dimensions to the loss, hoping that all frequencies are learned more equally. In a first attempt, data normalzed to zero mean and unit variance turned out to produce too small loss for training. With vanishing gradients the network could not learn at all. Therefore, in a second attempt the input data was normalized to zero mean and an equal variance of all dimensions of 10. With this specifications, the network was able to learn as well as without normalization.\\
As intended, the normalization helped to learn higher frequencies. The generated results with this setting contained more different frequencies from the range $[0,360]$. But the generated audio still converged to a constant distribution of frequencies after a few seconds. 
\begin{figure}[tbp]
  \centering
  \includegraphics[width=\columnwidth]{figures/original_frequencies.png}
  \caption{Frequencies of a piano piece used for training}
  \label{fig:original_frequencies}
\end{figure}
\begin{figure}[tbp]
  \centering
  \includegraphics[width=\columnwidth]{figures/generated_frequencies_20s_2.png}
  \caption{Frequencies of the generated audio piece}
  \label{fig:generated_frequencies}
\end{figure}
